---
title: "Homework 4"
author: "Qianlang Chen"

# CS 3130
# M 03/23/20

output: pdf_document

header-includes:
  - \usepackage{setspace}
  - \linespread{1.5}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

According to the properties of probability mass functions, the total probabality of all possible cases combined should equal to 1. We can form the following equation using this feature:
$$
\begin{aligned}
                                                                           \sum_{i=1}^4 p_i &= 1\\
                                                          \sum_{i=1}^4 \alpha \cdot (i + 1) &= 1\\
  \alpha \cdot (1 + 1) + \alpha \cdot (2 + 1) + \alpha \cdot (3 + 1) + \alpha \cdot (4 + 1) &= 1\\
                                                                                   14\alpha &= 1\\
                                                                                     \alpha &= \boxed{\frac{1}{14}}
\end{aligned}
$$

\pagebreak
# Problem 2

First, let us find the probability density function of $X$:
$$
f(x) = \begin{cases}
  0, &x < 0\text{ or }x > 2\\
  \dfrac{1}{b - a} = \dfrac{1}{2-0} = \dfrac{1}{2}, &0 \le x \le 2
\end{cases}
$$

According to the *Change-of-variable formula*, we have
$$
\begin{aligned}
  \mathbb{E}(g(X)) &= \int_{-\infty}^{\infty} g(x)\ f(x)\ \text{d}x\\
                   &= \int_{-\infty}^0 (x^3 + x^2) \cdot 0\ \text{d}x + \int_0^2 (x^3 + x^2) \cdot \frac{1}{2}\ \text{d}x + \int_2^{\infty} (x^3 + x^2) \cdot 0\ \text{d}x\\
                   &= 0 + \frac{10}{3} + 0\\
                   &= \boxed{\frac{10}{3}}
\end{aligned}
$$

\pagebreak
# Problem 3

## Part (a)

$$
P(X = k) = \begin{cases}
  0, &k \le 0\\
  (1 - \dfrac{1}{2})^{k-1} \cdot (\dfrac{1}{2})^1 = \dfrac{1}{2^k}, &k > 0
\end{cases}
$$

## Part (b)

$$
\begin{aligned}
  \mathbb{E}(X) &= \sum_k k \cdot P(X = k)\\
                &= \sum_{k=1}^{\infty} k \cdot \frac{1}{2^k}\\
                &= \frac{1}{2} + \frac{2}{4} + \frac{3}{8} + \cdots\\
                &= \boxed{2}
\end{aligned}
$$

\pagebreak
# Problem 4

**Probability Density Functions**

PDF of $X$:
$$
f(x) = \begin{cases}
  0, &x < -1\text{ or }x > 1\\
  \dfrac{1}{b - a} = \dfrac{1}{1-(-1)} = \dfrac{1}{2}, &-1 \le x \le 1
\end{cases}
$$

PDF of $Y$:
$$
f(y) = \begin{cases}
  0, &y < -4\text{ or }y > 4\\
  \dfrac{1}{b - a} = \dfrac{1}{4-(-4)} = \dfrac{1}{8}, &-4 \le y \le 4
\end{cases}
$$

PDF of $Z$:
$$
f(z) = \begin{cases}
  0, &z < 4\text{ or }z > 6\\
  \dfrac{1}{b - a} = \dfrac{1}{6-4} = \dfrac{1}{2}, &4 \le z \le 6
\end{cases}
$$

**Expected Values**

Expected Value of $X$:
$$\mathbb{E}(X) = \int_{-\infty}^{\infty} x \cdot f(x)\ \text{d}x = \int_{-1}^1 x \cdot \frac{1}{2}\ \text{d}x = 0$$

Expected Value of $Y$:
$$\mathbb{E}(Y) = \int_{-\infty}^{\infty} y \cdot f(y)\ \text{d}y = \int_{-4}^4 y \cdot \frac{1}{8}\ \text{d}y = 0$$

Expected Value of $Z$:
$$\mathbb{E}(Z) = \int_{-\infty}^{\infty} z \cdot f(z)\ \text{d}z = \int_4^6 z \cdot \frac{1}{2}\ \text{d}z = 5$$

**Variances**

Let us use the shortcut formula described in the textbook:
$$\text{Var}(X) = \mathbb{E}(X^2) - \mathbb{E}^2(X)$$

Variance of $X$:
$$
\begin{aligned}
  \text{Var}(X) &= \mathbb{E}(X^2) - \mathbb{E}^2(X)\\
                &= \int_{-1}^1 x^2 \cdot \frac{1}{2}\ \text{d}x - 0^2\\
                &= \frac{1}{3} - 0\\
                &= \boxed{\frac{1}{3}}
\end{aligned}
$$

Variance of $Y$:
$$
\begin{aligned}
  \text{Var}(Y) &= \mathbb{E}(Y^2) - \mathbb{E}^2(Y)\\
                &= \int_{-4}^4 y^2 \cdot \frac{1}{8}\ \text{d}y - 0^2\\
                &= \frac{16}{3} - 0\\
                &= \boxed{\frac{16}{3}}
\end{aligned}
$$

Variance of $Z$:
$$
\begin{aligned}
  \text{Var}(Z) &= \mathbb{E}(Z^2) - \mathbb{E}^2(Z)\\
                &= \int_4^6 z^2 \cdot \frac{1}{2}\ \text{d}z - 5^2\\
                &= \frac{76}{3} - 25\\
                &= \boxed{\frac{1}{3}}
\end{aligned}
$$

**Observations**

The variances of $X$ and $Z$ are the same because their random variables come from regions with the same size (2) and are both uniform.

\pagebreak
# Problem 5

## Part (a)

```{r}
# histogram the distribution of 10 thousand random gaussian samples
hist(rnorm(10000, 0, 1))

# the plot of the probability density function for the above random varibale
plot(dnorm(seq(-4, 4, length = 10000)))
```

## Part (b)

```{r}
# calculates the expected values of (x^4) empirically out of n samples
calc.e.x4 = function (n)
{
  x = 0
  for (i in 1:n)
    x = x + rnorm(1, 0, 1)^4
  
  return (x / n)
}

# 10 thousand samples
print(calc.e.x4(10000))

# 50 thousand samples
print(calc.e.x4(50000))
```

